---
title: "Data split and training a machine learning model"
author: "Steff Taelman"
date: "3/22/2023"
output: html_document
---

## loading in packages
You may need to install these first using `install.packages("package_name")`

```{r packages}
suppressPackageStartupMessages(
  c(library(dplyr), library(randomForest), library(caret))
  )
```



## Objective

This markdown file reads in the lung disease patient data and trains and tests a basic machine learning model on it. To ensure reproducibility across devices, please set the working directory to the parent directory of this script and the 'data' file (in my case `privacy_friendly_ML`):

```{r input paths, echo=FALSE}
# set the work directory
knitr::opts_knit$set(
  root.dir = '/Users/stefftaelman/Documents/Projects/privacy_friendly_ML', 
  echo=TRUE
  )
message(paste0("The current working directory is set to: ", getwd()))

# check the file names and paths
data_path = "./data/cancer_patient_data_sets.csv"
```


## Reading in the data

```{r data read in}
full_data <- read.csv(data_path, row.names = 1)
full_data$Level <- as.factor(full_data$Level)
```


## Some exploration
To frame the dataset in our text, we could do some basic data exploration here...

```{r data exploration}

```


## Splitting the data
Because ...

```{r splitting}
# set a seed to make our random split reproducible
set.seed(42)

# random 80-20 split
train <- full_data %>% dplyr::sample_frac(0.8)
test  <- dplyr::anti_join(full_data, train, by = 'Patient.Id')

# remove the patient ID label
train <- train[, -1]
test <- test[, -1]
```


## Train a basic machine learning model

```{r basic rf}
rf <- randomForest(
  Level ~ ., 
  data = train, 
  importance = TRUE,
  proximity = TRUE
  )
print(rf)
```

## Make predictions on the test data & check the accuracy

```{r}
predictions <- predict(rf, newdata = test)
confusionMatrix(predictions, test$Level)
```

