---
title: "Data split and training a machine learning model"
author: "Steff Taelman"
date: "3/22/2023"
output: html_document
---

## loading in packages
You may need to install these first using `install.packages("package_name")`

```{r packages}
suppressPackageStartupMessages(
  c(library(dplyr), library(randomForest), library(caret))
  )
```



## Objective

This markdown file reads in the lung disease patient data and trains and tests a basic machine learning model on it. To ensure reproducibility across devices, please set the working directory to the parent directory of this script and the 'data' file (in my case `privacy_friendly_ML`):

```{r input paths, echo=FALSE}
# set the work directory
knitr::opts_knit$set(
  root.dir = '/Users/stefftaelman/Documents/Projects/privacy_friendly_ML', 
  echo=TRUE
  )
message(paste0("The current working directory is set to: ", getwd()))

# check the file names and paths
data_path = "./data/cancer_patient_data_sets.csv"
```


## Reading in the data

```{r data read in}
full_data <- read.csv(data_path, row.names = 1)
full_data$Level <- as.factor(full_data$Level)
```


## Some exploration
To frame the dataset in our text, we could do some basic data exploration here, like setting out the age distribution of the patients, how many male/female, etc.

```{r data exploration}

```


## Splitting the data
Because we will eventually want to see how well our model is at predicting level of lung disease, we will randomly split the 1000 patients into two sets. We'll build our model based on the first one (which includes 700 patients), and have it make predictions for the remaining 300. At that point, we can then check how accurate these predictions are.

```{r splitting}
# set a seed to make our random split reproducible
set.seed(42)

# random 70-30 split
train <- full_data %>% dplyr::sample_frac(0.7)
test  <- dplyr::anti_join(full_data, train, by = 'Patient.Id')

# remove the patient ID label
train <- train[, -1]
test <- test[, -1]
```


## Train a basic machine learning model
We will now train a model to predict the link between the features (e.g. age, gender, smoking) and the label (level of lung disease). By modelling this link, we can make the model predict the label for new individuals for which we do not yet know their level of lung disease.

```{r basic rf}
rf <- randomForest(
  Level ~ ., 
  data = train, 
  ntree = 5,
  importance = TRUE,
  proximity = TRUE
  )
print(rf)
```

Let's take a quick look at what 'features' the model uses to make its predictions:

```{r}
varImpPlot(rf)
```

It seems the most telling factors are whether or not someone snores, their level of fatigue, and whether they are a passive smoker.

## Make predictions on the test data & check the accuracy

```{r}
predictions <- predict(rf, newdata = test)
confusionMatrix(predictions, test$Level)
```
All predictions are correct!
